New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions

### 2022-08-19 paper/appendix.Rmd
first pass at my changes

---
title: |
  Appendix: New and simplified manual controls for projection and slice tours
type: Short Technical Note
author:
  - name: Alex Aumann
    affil: a
    email: aaum0002@student.monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
  - name: Ursula Laa
    affil: b
    email: ursula.laa@boku.ac.at
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \usepackage{graphicx}
  \usepackage{nicematrix}
  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
output: rticles::tf_article
keep_tex: true
---

@@

For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$. Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$. Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
\end{enumerate}
\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
\end{enumerate}
The random choice for the components of $A$ not in the projection matrix allows the exploration of dimensions perpendicular to the projection plane.

@@

@@

The following functions are implemented in the Mathematica package `mmtour.wl`:

- `ProjectionPlot[data, projmat, xRange(=Automatic), yRange(=Automatic), 
colour(=Automatic)]` with arguments

    - `data`: A list of data matrices (they must all have the same dimensionality)
    - ``projmat`: A projection matrix which describes the projection plane. This is the 
initial projection and does not need to be orthonormal (XXX is this right?).
    - ``xRange`: A list which details the range of the horizontal axis.
    - ``yRange`: A list which details the range of the vertical axis.
    - ``colour`: A list of graphics directives to be applied to each data matrix.

There are problems with the default values for x and y ranges and colour. It would be 
better to fix the ranges and colour based on the data as well as the origin of axes 
and the size of the points.

**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 1**

The format for the printed projection matrix can be improved also. Maybe fix the 
number of digits and the position of the label.

- `SlicePlot[data, projmat, centrePoint, height, heightRange, ptSize1(=0.005), ptSize2(=0.004), minDist(=0)]` with arguments (used to be called SliceDynamic in Alex's first version, the name SliceDynamic was taken by something else below)

    - `data`: A data matrix
    - `projmat`: The inital projection matrix
    - `centrepoint`: The initial position of the slice
    - `height`: The initial height of the slice
    - `heightRange`: The range of heights
    - `ptSize1`: The size of the points that exist within the slice
    - `ptSize2`: The size of the points that exist outside of the slice

This function requires some auxiliary functions. It does not accept more than one data set.

**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data?**

It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes

- `SliceDynamic[data, projmat, centrePoint, height, heightRange, ptSize1 : 0.005, ptSize2 : 0.004]` with arguments

- `data`: A data matrix with an extra grouping or clustering column (numerical) 
- `projmat`: The initial projection matrix
- `centrepoint`: The initial position of the slice
- `height`: The initial height of the slice
- `heightRange`: The range of heights
- `ptSize1`: The size of the points that exist within the slice
- `ptSize2`: The size of the points that exist outside of the slice

**Does this need to be a separate function from SlicePlot? Fix 2**

**It would be good if the initial projection matrix (and maybe the center) had default values. Perhaps a random matrix with the correct dimensions read from the data? Fix 3**

**It may be good to fix the size of the points, the center of the axes and the scales of the axes rather than allow them to change as the projection and slicing changes. Fix 4**

**I have found it useful to input different data sets rather than have this function sort them automatically using the last column. One data set with a flag column at the end, is best! Is the code too complicated? Fix 5**

- `Projected2DSliderPlot[]`

This function uses a different display for the projection that I find very useful. However as it stands it is specific to one data set. Needs to be generalised to arbitrary data if we want to keep it.

**For Alex to do**

- **We would need a screenshot (and video) for each of these.**
- **Fix 1, 2, 3, 4, 5**
- **Code consistent for applications to work - see slice_tour_compare_sets.nb. See below**
    - **We need to make sure the drawing range is fixed (no resizing of the axis when changing projection or slicing). It could be nice if there was a manual zoom in/zoom out to change scale occasionally.**
    - **Should also make sure the size of the points is fixed (maybe this will be automatic if the drawing range is fixed?), bigger points would be preferred, currently the size is sometimes really tiny and it becomes difficult to see anything.**
    - **Can we have some heuristic to decide the range for the slice thickness slider? In the second graph there is a tiny part of the slider where the changes are interesting, difficult to navigate since this makes it super sensitive to any small movement, while most of the options are not interesting. My suggestion would be to keep (as I understand is the case now) the maximum thickness (all points in) as the upper bound, but instead of zero use a thickness below which slices tend to be empty as the lower bound.**
    - **For the olives example I think the relation between x2 and the blue region could be interesting.**
    - **pdfsense data: can we use random samples instead of the first 10k rows? Then we should still be able to see the bigger picture (with picking the first ones we get some artificial structure from how the data was sampled initially).**
    - **I was playing around with this today. Looks like the axes are fixed when changing slice thickness, but not when rotating?**
    - **Anyways, some observations that might be useful, but all of this seems difficult for including in a paper:**
    - **In the first example we can show how the yellow group is associated with the first variable (x1, I guess this is C9?) - when slicing through the center it only appears when x1 has a bigger component on the projection, it is associated with low values of x1. Since they are not near the average value of x1 the points do not get captured in a slice where x1 is not important in the projection. We can move the center point to -1 in the first component to then explore how the yellow group relates to the other three variables. Much less clear, maybe some indication of smaller x2 and larger x3 for this.** 
    - **The second example is too busy for manual exploration I would say (too many axes to move around separately).**
    - **Third example could be similar to the first: blue region requires small values of x2, can be understood from the exploration, then move center point to -0.5 in x2 to explore further. Now interpretation becomes tricky. Sometimes blue and yellow seem just randomly everywhere, sometimes blue more contained in an ellipsoid around the center (indicating some correlation between variables). For example try zeroing out components 2 and 3 to see this slice in x1 vs x4.**
    - **For the last example x4 seems to be the one to work with. I feel like this is again very similar, maybe better to instead explore other examples (maybe something without slicing?)**

>>>>

---
title: |
  Appendix: New and simplified manual controls for projection and slice tours
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a
    email: ursula.laa@boku.ac.at
  - name: Alex Aumann
    affil: b
    email: aaum0002@student.monash.edu
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
  - name: German Valencia
    affil: b
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      School of Physics and Astronomy, Monash University
  - num: c
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \usepackage{graphicx}
  \usepackage{nicematrix}
  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
output: rticles::tf_article
keep_tex: true
---

@@

For $d=2$ projections, the projection matrix is the sub-matrix of $A$ formed by its first two columns. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots,}$, orthonormality of the projection matrix is expressed as $P_i\cdot P_j=\delta_{ij}, {i,j=1,2}$. Movement of the cursor takes the two components ${x_{m1},x_{m2}}$ into a selected new value ${a,b}$. Although the motion is constrained by $a^2+b^2\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${x_{m1},x_{m2}}\to {a,b}$
\item The freedom to change the components $A_{i3}\cdots A_{ip}$ (the columns of $A$ not corresponding to the projection matrix) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $A_{m3}=\sqrt{1-a^2-b^2},~A_{m,k>3}=0$
\item For other rows, $A_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
\end{enumerate}
\item this results in the orthonormal basis $A^*$ and a new projection matrix with $P_{m1}=a,~P_{m2}=b$.
\end{enumerate}
The random completion of $A$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane. However, it renders projections in a way that is not continuouns and may be distracting. This can be alleviated by replacing the random completion with a rule restricting the size of the jumps.

@@

{\textcolor{blue} Do we need three methods? is this one implemented somewhere?}

@@

Here we describe the functions implemented in the Mathematica package `mmtour.wl`. The main function is

- `SliceDynamic[data, projmat, height, heightRange, legendQ(=1), flagQ(=1), colorFunc(=ColorData[97])]` with arguments

- `data`: A data matrix with an extra grouping that has a label (string) and a flag  (numerical) 
- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
- `height`: The initial height of the slice.
- `heightRange`: The range of slice heights to be explored.
- `legendQ`: A flag signaling whether the data matrix includes a column with group labels (1) or not (0).
- `flagQ`: A flag signaling whether the data matrix includes a numerical flag for the groups.
- `colorFunc`: specifies the colour mapping for the different groups.

The function renders a manual tour with controls that allow the user to navigate through different projections. For a given projection, a slider permits the user to change the slide width within its range and a separate box switches the view to a projection. The user can also change the centre point of slices; the size of the points rendered and the scale of the plot through sliders. An additional box displays the coordinates of the projection matrix explicitly.

This function can be used for ungrouped data setting  legendQ and flagQ to 0.

To remove a group from the visual display, the user can choose specific colours for the entry colorFunc. For example, {Red,White,Green} would display groups 1 (red) and 3 (green) while making 2 invisible.

Other functions are:

- `ProjectionPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]` with arguments

- `data`: A data matrix with one or more groups labeled by a flag  (numerical) 
- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
- `legendNames`: an optional list of labels for the groups.
- `colorFunc`: specifies the colour mapping for the different groups.

This function is similar to SliceDynamic but can only display projections. Using it instead of SliceDynamic simplifies the display and is more efficient. The data file should not contain labels for groups, if these are desired they are passed through the optional argument legendNames as {"group 1", "group 2",...}, for example.

The display allows the user to navigate through different projections, choose the point size and zoom the factor, and explicitly display the coordinates of the projection.

- `ProjectedLocatorPlot[data, projmat, legendNames(={}), colorFunc(=ColorData[97])]` with arguments

- `data`: A data matrix with one or more groups labeled by a flag  (numerical) 
- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
- `legendNames`: an optional list of labels for the groups.
- `colorFunc`: specifies the colour mapping for the different groups.

This function is very similar to ProjectionPlot but uses a different display for navigation of  projections in which each coordinate is represented on a separate dial.

- `VisualiseSliceDynamic[data, projmat, height, heightRange]` with arguments

- `data`: A data matrix containg only one set.In this case the data matrix should not contain column labels.
- `projmat`: The initial projection matrix. It is possible to start with a random projection with the input "random" (including the quotes) in this entry field.
- `height`: The initial height of the slice.
- `heightRange`: The range of slice heights to be explored.

This function differs from SliceDynamic in that it looks at only one data set but displays both the points inside and outside the slice. The display panel allows the user to specify the size of both sets of points (inside and outside) along with the slice height, centre point and zoom level.

### 2022-08-19 paper/appendix.tex
first pass at my changes

\author{\name{Alex Aumann$^{a}$, German Valencia$^{a}$, Ursula
Laa$^{b}$, Dianne Cook$^{c}$}
\affil{$^{a}$School of Physics and Astronomy, Monash
University; $^{b}$Institute of Statistics, University of Natural
Resources and Life Sciences, Vienna; $^{b}$Department of Econometrics
and Business Statistics, Monash University}
}

\thanks{CONTACT Alex
Aumann. Email: \href{mailto:aaum0002@student.monash.edu}{\nolinkurl{aaum0002@student.monash.edu}}, German
Valencia. Email: \href{mailto:german.valencia@monash.edu}{\nolinkurl{german.valencia@monash.edu}}, Ursula
Laa. Email: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}}

@@

The random choice for the components of \(A\) not in the projection
matrix allows the exploration of dimensions perpendicular to the
projection plane.

@@

@@

The following functions are implemented in the Mathematica package
\texttt{mmtour.wl}:

\begin{itemize}
\item
  \texttt{ProjectionPlot{[}data,\ projmat,\ xRange(=Automatic),\ yRange(=Automatic),\ \ colour(=Automatic){]}}
  with arguments

  \begin{itemize}
  \tightlist
  \item
    \texttt{data}: A list of data matrices (they must all have the same
    dimensionality)
  \item
    `\texttt{projmat}: A projection matrix which describes the
    projection plane. This is the initial projection and does not need
    to be orthonormal (XXX is this right?).
  \item
    `\texttt{xRange}: A list which details the range of the horizontal
    axis.
  \item
    `\texttt{yRange}: A list which details the range of the vertical
    axis.
  \item
    `\texttt{colour}: A list of graphics directives to be applied to
    each data matrix.
  \end{itemize}
\end{itemize}

There are problems with the default values for x and y ranges and
colour. It would be better to fix the ranges and colour based on the
data as well as the origin of axes and the size of the points.

\textbf{It would be good if the initial projection matrix (and maybe the
center) had default values. Perhaps a random matrix with the correct
dimensions read from the data? Fix 1}

The format for the printed projection matrix can be improved also. Maybe
fix the number of digits and the position of the label.

\begin{itemize}
\item
  \texttt{SlicePlot{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1(=0.005),\ ptSize2(=0.004),\ minDist(=0){]}}
  with arguments (used to be called SliceDynamic in Alex's first
  version, the name SliceDynamic was taken by something else below)

  \begin{itemize}
  \tightlist
  \item
    \texttt{data}: A data matrix
  \item
    \texttt{projmat}: The inital projection matrix
  \item
    \texttt{centrepoint}: The initial position of the slice
  \item
    \texttt{height}: The initial height of the slice
  \item
    \texttt{heightRange}: The range of heights
  \item
    \texttt{ptSize1}: The size of the points that exist within the slice
  \item
    \texttt{ptSize2}: The size of the points that exist outside of the
    slice
  \end{itemize}
\end{itemize}

This function requires some auxiliary functions. It does not accept more
than one data set.

\textbf{It would be good if the initial projection matrix (and maybe the
center) had default values. Perhaps a random matrix with the correct
dimensions read from the data?}

It may be good to fix the size of the points, the center of the axes and
the scales of the axes rather than allow them to change as the
projection and slicing changes

\begin{itemize}
\item
  \texttt{SliceDynamic{[}data,\ projmat,\ centrePoint,\ height,\ heightRange,\ ptSize1\ :\ 0.005,\ ptSize2\ :\ 0.004{]}}
  with arguments
\item
  \texttt{data}: A data matrix with an extra grouping or clustering
  column (numerical)
\item
  \texttt{projmat}: The initial projection matrix
\item
  \texttt{centrepoint}: The initial position of the slice
\item
  \texttt{height}: The initial height of the slice
\item
  \texttt{heightRange}: The range of heights
\item
  \texttt{ptSize1}: The size of the points that exist within the slice
\item
  \texttt{ptSize2}: The size of the points that exist outside of the
  slice
\end{itemize}

\textbf{Does this need to be a separate function from SlicePlot? Fix 2}

\textbf{It would be good if the initial projection matrix (and maybe the
center) had default values. Perhaps a random matrix with the correct
dimensions read from the data? Fix 3}

\textbf{It may be good to fix the size of the points, the center of the
axes and the scales of the axes rather than allow them to change as the
projection and slicing changes. Fix 4}

\textbf{I have found it useful to input different data sets rather than
have this function sort them automatically using the last column. One
data set with a flag column at the end, is best! Is the code too
complicated? Fix 5}

\begin{itemize}
\tightlist
\item
  \texttt{Projected2DSliderPlot{[}{]}}
\end{itemize}

This function uses a different display for the projection that I find
very useful. However as it stands it is specific to one data set. Needs
to be generalised to arbitrary data if we want to keep it.

\textbf{For Alex to do}

\begin{itemize}
\tightlist
\item
  \textbf{We would need a screenshot (and video) for each of these.}
\item
  \textbf{Fix 1, 2, 3, 4, 5}
\item
  \textbf{Code consistent for applications to work - see
  slice\_tour\_compare\_sets.nb. See below}

  \begin{itemize}
  \tightlist
  \item
    \textbf{We need to make sure the drawing range is fixed (no resizing
    of the axis when changing projection or slicing). It could be nice
    if there was a manual zoom in/zoom out to change scale
    occasionally.}
  \item
    \textbf{Should also make sure the size of the points is fixed (maybe
    this will be automatic if the drawing range is fixed?), bigger
    points would be preferred, currently the size is sometimes really
    tiny and it becomes difficult to see anything.}
  \item
    \textbf{Can we have some heuristic to decide the range for the slice
    thickness slider? In the second graph there is a tiny part of the
    slider where the changes are interesting, difficult to navigate
    since this makes it super sensitive to any small movement, while
    most of the options are not interesting. My suggestion would be to
    keep (as I understand is the case now) the maximum thickness (all
    points in) as the upper bound, but instead of zero use a thickness
    below which slices tend to be empty as the lower bound.}
  \item
    \textbf{For the olives example I think the relation between x2 and
    the blue region could be interesting.}
  \item
    \textbf{pdfsense data: can we use random samples instead of the
    first 10k rows? Then we should still be able to see the bigger
    picture (with picking the first ones we get some artificial
    structure from how the data was sampled initially).}
  \item
    \textbf{I was playing around with this today. Looks like the axes
    are fixed when changing slice thickness, but not when rotating?}
  \item
    \textbf{Anyways, some observations that might be useful, but all of
    this seems difficult for including in a paper:}
  \item
    \textbf{In the first example we can show how the yellow group is
    associated with the first variable (x1, I guess this is C9?) - when
    slicing through the center it only appears when x1 has a bigger
    component on the projection, it is associated with low values of x1.
    Since they are not near the average value of x1 the points do not
    get captured in a slice where x1 is not important in the projection.
    We can move the center point to -1 in the first component to then
    explore how the yellow group relates to the other three variables.
    Much less clear, maybe some indication of smaller x2 and larger x3
    for this.}
  \item
    \textbf{The second example is too busy for manual exploration I
    would say (too many axes to move around separately).}
  \item
    \textbf{Third example could be similar to the first: blue region
    requires small values of x2, can be understood from the exploration,
    then move center point to -0.5 in x2 to explore further. Now
    interpretation becomes tricky. Sometimes blue and yellow seem just
    randomly everywhere, sometimes blue more contained in an ellipsoid
    around the center (indicating some correlation between variables).
    For example try zeroing out components 2 and 3 to see this slice in
    x1 vs x4.}
  \item
    \textbf{For the last example x4 seems to be the one to work with. I
    feel like this is again very similar, maybe better to instead
    explore other examples (maybe something without slicing?)}
  \end{itemize}
\end{itemize}

>>>>

\author{\name{Ursula Laa$^{a}$, Alex Aumann$^{b}$, Dianne
Cook$^{c}$, German Valencia$^{b}$}
\affil{$^{a}$Institute of Statistics, University of Natural Resources
and Life Sciences, Vienna; $^{b}$School of Physics and Astronomy, Monash
University; $^{c}$Department of Econometrics and Business Statistics,
Monash University}
}

\thanks{CONTACT Ursula
Laa. Email: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}, Alex
Aumann. Email: \href{mailto:aaum0002@student.monash.edu}{\nolinkurl{aaum0002@student.monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, German
Valencia. Email: \href{mailto:german.valencia@monash.edu}{\nolinkurl{german.valencia@monash.edu}}}

@@

The random completion of \(A\) outside the projection matrix provides an
exploration of dimensions orthogonal to the projection plane. However,
it renders projections in a way that is not continuouns and may be
distracting. This can be alleviated by replacing the random completion
with a rule restricting the size of the jumps.

@@

\{\textcolor{blue} Do we need three methods? is this one implemented
somewhere?\}

@@

Here we describe the functions implemented in the Mathematica package
\texttt{mmtour.wl}. The main function is

\begin{itemize}
\item
  \texttt{SliceDynamic{[}data,\ projmat,\ height,\ heightRange,\ legendQ(=1),\ flagQ(=1),\ colorFunc(=ColorData{[}97{]}){]}}
  with arguments
\item
  \texttt{data}: A data matrix with an extra grouping that has a label
  (string) and a flag (numerical)
\item
  \texttt{projmat}: The initial projection matrix. It is possible to
  start with a random projection with the input ``random'' (including
  the quotes) in this entry field.
\item
  \texttt{height}: The initial height of the slice.
\item
  \texttt{heightRange}: The range of slice heights to be explored.
\item
  \texttt{legendQ}: A flag signaling whether the data matrix includes a
  column with group labels (1) or not (0).
\item
  \texttt{flagQ}: A flag signaling whether the data matrix includes a
  numerical flag for the groups.
\item
  \texttt{colorFunc}: specifies the colour mapping for the different
  groups.
\end{itemize}

The function renders a manual tour with controls that allow the user to
navigate through different projections. For a given projection, a slider
permits the user to change the slide width within its range and a
separate box switches the view to a projection. The user can also change
the centre point of slices; the size of the points rendered and the
scale of the plot through sliders. An additional box displays the
coordinates of the projection matrix explicitly.

This function can be used for ungrouped data setting legendQ and flagQ
to 0.

To remove a group from the visual display, the user can choose specific
colours for the entry colorFunc. For example, \{Red,White,Green\} would
display groups 1 (red) and 3 (green) while making 2 invisible.

Other functions are:

\begin{itemize}
\item
  \texttt{ProjectionPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
  with arguments
\item
  \texttt{data}: A data matrix with one or more groups labeled by a flag
  (numerical)
\item
  \texttt{projmat}: The initial projection matrix. It is possible to
  start with a random projection with the input ``random'' (including
  the quotes) in this entry field.
\item
  \texttt{legendNames}: an optional list of labels for the groups.
\item
  \texttt{colorFunc}: specifies the colour mapping for the different
  groups.
\end{itemize}

This function is similar to SliceDynamic but can only display
projections. Using it instead of SliceDynamic simplifies the display and
is more efficient. The data file should not contain labels for groups,
if these are desired they are passed through the optional argument
legendNames as \{``group 1'', ``group 2'',\ldots\}, for example.

The display allows the user to navigate through different projections,
choose the point size and zoom the factor, and explicitly display the
coordinates of the projection.

\begin{itemize}
\item
  \texttt{ProjectedLocatorPlot{[}data,\ projmat,\ legendNames(=\{\}),\ colorFunc(=ColorData{[}97{]}){]}}
  with arguments
\item
  \texttt{data}: A data matrix with one or more groups labeled by a flag
  (numerical)
\item
  \texttt{projmat}: The initial projection matrix. It is possible to
  start with a random projection with the input ``random'' (including
  the quotes) in this entry field.
\item
  \texttt{legendNames}: an optional list of labels for the groups.
\item
  \texttt{colorFunc}: specifies the colour mapping for the different
  groups.
\end{itemize}

This function is very similar to ProjectionPlot but uses a different
display for navigation of projections in which each coordinate is
represented on a separate dial.

\begin{itemize}
\item
  \texttt{VisualiseSliceDynamic{[}data,\ projmat,\ height,\ heightRange{]}}
  with arguments
\item
  \texttt{data}: A data matrix containg only one set.In this case the
  data matrix should not contain column labels.
\item
  \texttt{projmat}: The initial projection matrix. It is possible to
  start with a random projection with the input ``random'' (including
  the quotes) in this entry field.
\item
  \texttt{height}: The initial height of the slice.
\item
  \texttt{heightRange}: The range of slice heights to be explored.
\end{itemize}

This function differs from SliceDynamic in that it looks at only one
data set but displays both the points inside and outside the slice. The
display panel allows the user to specify the size of both sets of points
(inside and outside) along with the slice height, centre point and zoom
level.

### 2022-08-19 paper/paper.tex
mapping out discussion
discussion points
first pass at my changes

\begin{abstract}
Something here
\end{abstract}

@@

@@

@@

@@

@@

The authors gratefully acknowledge the support of the Australian
Research Council. The paper was written in \texttt{rmarkdown}
\citep{rmarkdown} using \texttt{knitr} \citep{knitr}. The scatterplot
matrix of the penguins data was produced by the GGally \citep{GGally}
package built on ggplot2 \citep{ggplot2} graphics.

>>>>

\begin{abstract}
We implement a new version of a manual tour that takes advantage of
existing features in Mathematica to construct functions that interact
with the position of the cursor. We demonstrate the usefulness of this
new tool with an example exploring region boundaries in classification
problems.
\end{abstract}

@@

\{\textcolor{blue}Another very useful feature is that the user can keep
the projection fixed while slicing across the orthogonal space and
changing the centre point used for slicing. This can serve to illustrate
how the orthogonal space affects correlations in a slice and how they
may change or disappear in a projection.\}

@@

XXX Equation for arbitrary center for slice; and move the ``the
implementation \ldots{}'' into the software section once that is
constructed; and ``potential adjustments for future'' into the section
on desirable controls

@@

Mathematica provides much utility and versatility, such as inbuilt data
visualisation, data manipulation and analysis, dynamic functionality,
and symbolic and numeric computation. Most of this inbuilt functionality
is quite user-friendly and can be found in the Mathematica
documentation; typically, numerous examples are provided within the
documentation, and some possible issues are outlined there. Also, it is
relatively simple to create dynamic objects via the inbuilt commands
Manipulate or DynamicModule. Control objects, such as sliders, locator
panes, and input fields, can then be used on dynamic variables, and when
there is a change in the dynamic variable the dynamic objects which
contain that dynamic variable will be updated. This is the essential
ingredient for this implementation of the manual tour. However, it can
be hard to alter the behaviour of the control objects such as setting
the 1D slider on a different scale (Slider increments add 10 points into
the slice rather than increasing the slice height by 0.1 for example).
\{\textcolor{red}Alex, is this still true with the controls you are
using now?\}

The examples presented in the penguins' notebooks all utilise our
primary new function, SliceDynamic. This function typically accepts
grouped data in the form of a matrix where the second last column
details the name of the group and the last column details the group
index. After specifying the initial slice height and the slice range,
the user is then presented with an interactive display in which the
control objects appear on the left and the slice visulisation appears on
the right. The user can change the orientation of the slice via the
locator pane, which changes the projection matrix; a slider controls the
slice height; and there is an input field, which changes the slice
centre. The user can also change the appearance of the plot by zooming
into the center or changing the point size with sliders provided. The
current projection can be displayed to contrast it with the slice by
ticking a box. The explicit projection matrix can be displayed via
another checkbox. The Slice Tour* notebooks include the additional
functions: ProjectionPlot, Projected2DSliderPlot, and
VisualiseSliceDyanmic. ProjectionPlot is very similar to SliceDynamic,
except it only displays projections and VisualiseSliceDynamic allows you
to discern which points exist inside the slice and outside it. However,
ProjectedLocatorPlot displays the interactive slightly differently.
Several locator panes are displayed above the plot and each one
corresponds to a row in the projection matrix but behaves the same as
ProjectionPlot. \{\textcolor{blue}(this may be easier to implement in
another language (?) )\}

@@

Recap contribution

Other applications

Where else might it be useful to use mathematica

Maybe extensions for slicing

@@

The authors gratefully acknowledge the support of the Australian
Research Council and the ResearchFirst undergraduate research program at
Monash University. The paper was written in \texttt{rmarkdown}
\citep{rmarkdown} using \texttt{knitr} \citep{knitr}. The scatterplot
matrix of the penguins data was produced by the GGally \citep{GGally}
package built on ggplot2 \citep{ggplot2} graphics. We thank the
Institute of Statistics, BOKU, for their hospitality while part of this
work was conducted.
